CODE CHECK

### 2020-01-31 codecheck-paper.Rmd
Update codecheck-paper.Rmd

- **Principles**: https://codecheck.org.uk/
- relation to "proper" citation of data and software, and depositing of data and software in suitable repos _besides_ the bundle for CODE CHECK: do it, only provide the concrete analysis script for the check
- **Why is it useful?**
  - reduce the barrier to evaluating non-text parts of research
  - acknowledges different skill sets and existing high load on reviewers
  - CC breaks problem that publishers staff expertise today is not likely to suffice
  - engaging ECRs is a perfet fit (they are up-to-date in methods, are introduced to peer review)
  - relation to research compendia? point it out for authors unsure about how to structure their workflow
  - CrossRef and Publons and ORCID have features around reviews
  - shifting burden from reviewer to author (but not too much)
- keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e. a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines

@@

- need publishers build up data science experts who can conduct checks?
- integration of software/data citation checks benefitial?
- bot-support and tools (linter)
- Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?
- Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
- can CODE CHECK work for preprints? how?

>>>>

- **Principles**: https://codecheck.org.uk/
- relation to "proper" citation of data and software, and depositing of data and software in suitable repos _besides_ the bundle for CODE CHECK: do it, only provide the concrete analysis script for the check
- **Why is it useful?**
  - reduce the barrier to evaluating non-text parts of research
  - acknowledges different skill sets and existing high load on reviewers
  - CC breaks problem that publishers staff expertise today is not likely to suffice
  - engaging ECRs is a perfet fit (they are up-to-date in methods, are introduced to peer review)
  - relation to research compendia? point it out for authors unsure about how to structure their workflow
  - CrossRef and Publons and ORCID have features around reviews
  - shifting burden from reviewer to author (but not too much)
- keynote https://doi.org/10.7557/5.4963: in discussion, the speaker mentioned value of peer review for "baseline reassurance", i.e. a paper has at least been checked by someone with an understanding, to increase trust especially if looking at papers from other disciplines
- **report** must fulfil some requirements to make clear what is really checked and not mislead - readers will have very different understandings of a green checkmark!

@@

- need publishers build up data science experts who can conduct checks?
- integration of software/data citation checks benefitial?
- bot-support and tools (linter)
- Does nudging towards better practices work (highlight the successes) or is any check with opt-in doomed to fail?
- Do we need more than one codechecker? Is there a similar diversity as in a peer review where more opinions are needed?
- can CODE CHECK work for preprints? how? is a re-check required?
- risk: is the hurdle too low? are "levels" of checks are solution to increase transparency of check's level of detail
